---
title: "Estimation of the Value of $R_t$ by Municipio in each State for the Map"
author: "Joaquín Salas"
date: "June 7, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=3, fig.height=2)
```

## Introduction
This program estimates $R_t$ for the municipios of all the states in Mexico. 
We base our implementation in EpiEstim, a library package for **R** (the language) discussed 
in the article *A New Framework and Software to Estimate Time-Varying Reproduction*, 2013, and implemented in the library EpiEstim described in the link \url{https://cran.r-project.org/web/packages/EpiEstim/EpiEstim.pdf}

The authors of EpiEstim assumed that the rate at which a positive contagious person follows a Poisson process. Using a Gamma prior, the posterior results in a Gamma distribution. We use the expected value and standard deviation as the values of $R_t$ and its uncertainty. Following the recommendation, we do not compute $R_t$ when the number of cases is less than 12. Also, although estimated for influenza,  we use  a mean equal to 2.6 and a standard deviation equal to  1.5 for the parameters for the Gamma prior distribution.

To construct the graph, we compare the records between the most current dataset and the day before. We find the furthest date where there was a change from one day to the next, in the number of confirmed cases. We show a dashed line for the estimation of $R_t$ from then on.



```{r epidemiology}

#suppressMessages(require(ggmap)) # latitude-longitude
library(EpiEstim)
library(incidence) #incidence object
library(distcrete) #discretise a pdf
library(epitrix) #from mu, cv to scale and shape in the Gamma distribution
library(latex2exp) #include latex
library(ggplot2) #ggplot
source("readData.R")
source("estimateR0Value.R")


#asigna la llave API
#register_google(key = "AIzaSyDieTBLAqPlfc9kCShr4VUal-ou4w2ZbhU")


#working directories
code.dir = 'E:/Documents/informs/research/2020.05.18 prediction/code/'
data.dir = 'E:/Documents/informs/research/2020.05.18 prediction/data/'
setwd(code.dir)

#directories with files that  I use across many applications related to COVID
common.data.dir = 'E:\\Documents\\informs\\research\\covid-common\\data\\'
common.data = 'E:\\Documents\\informs\\research\\covid-common\\'



#files in the data.dir directory corresponding to the Health Ministery data
files.pattern = "*COVID19MEXICO.csv"
files <- list.files(path = common.data.dir, pattern =  files.pattern)


#############
#read the data from the last file of cases
filename = paste(common.data.dir, files[length(files)-1], sep = "")
covid.before = readData(filename)

filename = paste(common.data.dir, files[length(files)], sep = "")
covid = readData(filename)



#filter in positives
covid.pos = covid[covid$RESULTADO==1,]

#filter in positives
covid.pos.before = covid.before[covid.before$RESULTADO==1,]


```

These are the daily confirmed positives for COVID-19. The horizontal line indicates the date and the vertical one, the daily incidence.

```{r positives}





```

The following figure plots the estimated value of $R_t$ for the given epidemic curve. SI is the serial interval, the probability distribution for the period between the onset of the symptoms in the case and its infector. 

```{r value-of-R}



#read the list of states

filename = paste(common.data, "estados-lat-lon.csv", sep = "")
estados = read.csv(filename)


#read the list of municipios

filename = paste(common.data, "municipios-lat-lon.csv", sep = "")
municipios  = read.csv(filename)

#forza a realizar una predicción para el primer estado
old.clave.state = -1

i = 1
for (entry in seq(1, dim(municipios)[1])) {
  #print(entry)
  
  #extract the municipio name
  municipio.name = municipios$MUNICIPIO[entry]
  #print(droplevels(municipio.name))
  
  clave.municipio = municipios$CLAVE_MUNICIPIO[entry]
  clave.state = municipios$CLAVE_ENTIDAD[entry]
  
  if (!(clave.state  == old.clave.state)) {
    state.name = estados$ENTIDAD_FEDERATIVA[clave.state]
    print(droplevels(state.name))
    #obtain the cases for this state by state of residency 
    confirmed.state = covid.pos[covid.pos$ENTIDAD_RES ==  clave.state,]
    confirmed.state.before = covid.pos.before[covid.pos.before$ENTIDAD_RES ==  clave.state,]
    #estimate R0 for this municipio
    estimate = estimateR0Value(confirmed.state,  
                               confirmed.state.before, state.name,1)
    
    #fill up the results 
    res =  
      data.frame(estado = state.name, 
                 Rt = estimate$Rt, 
                 sd = estimate$sd, 
                 fecha = estimate$fecha, 
                 positivos = estimate$positivos, 
                 latitud = estados$LATITUD[clave.state], 
                 longitud = estados$LONGITUD[clave.state])
    #accumulate the results in the data frame
    if (i == 1){
      df.state = res
    } else {
      
      df.state = rbind(df.state, res)
      
    }
    
    old.clave.state = clave.state
  }
  
  
  
  #obtain the positives for this municipio out of the state positives
   confirmed.municipio = covid.pos[covid.pos$MUNICIPIO_RES ==  clave.municipio &
                                      covid.pos$ENTIDAD_RES == clave.state,]
    confirmed.municipio.before = 
      covid.pos.before[covid.pos.before$MUNICIPIO_RES ==  clave.municipio &
                         covid.pos.before$ENTIDAD_RES == clave.state,]
  
  
  
  
  
  #estimate R0 for this municipio
  estimate = estimateR0Value(confirmed.municipio,  
                             confirmed.municipio.before, municipio.name,1)
  
  
  #fill up the results 
  res =  
    data.frame(estado = state.name, 
               municipio = municipio.name, 
               Rt = estimate$Rt, 
               sd = estimate$sd,
               fecha = estimate$fecha, 
               positivos = estimate$positivos, 
               latitud = municipios$LATITUD[entry], 
               longitud = municipios$LONGITUD[entry])
   
                 
  #accumulate the results in the data frame
  if (i == 1){
    df.municipio = res
  } else {
  
    df.municipio = rbind(df.municipio, res)
  
  }
  
  i = i + 1
}



filename = paste(data.dir, "state-municipio-Rt-lat-lon3.csv", sep = "")
write.csv(df.municipio, filename)

filename = paste(data.dir, "state-Rt-lat-lon3.csv", sep = "")
write.csv(df.state, filename)

```


```{r pressure, echo=FALSE}
